---
title: "Final project"
date: "The Date"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, include = TRUE)
```

## Data 

  * log (for progress tracking purpose)
  
```{r}
project = list()
```
  


```{r}
# import data
library(readr)
enrollmentRate_department <- read_csv("data/學12-1.新生註冊率-以「系所」統計.csv")

# log 
project$data$enrollmentRate_department = list(
  value=enrollmentRate_department,
  description="enrollmentRate_department",
  source="https://data.gov.tw/dataset/26228"
)
```


### Checking

  * rename
  
```{r}
names(enrollmentRate_department)[c(1,2,5,7,8,10:14)]

names(enrollmentRate_department)[c(1,2,5,7,8,10:14)] <-  # notice 10:14
  c("schoolYear", "typeByFundingSource","school", "department","day_night", 
  "freshmenFresh","freshmenOld","freshmenFreshRegistered",
  "foreignStudentsRegistered", "netFreshmenRegistrationRate")

dfPartial = enrollmentRate_department[c(1,2,5,7,8,10:14)]
```

```{r}
dplyr::glimpse(dfPartial)
```

  * wrong class?

```{r}
library(magrittr)
dplyr::glimpse(dfPartial)
names(dfPartial)
```


```{r correct classes}
# elements 2 3 4 5 should be factor
factorPos = c(2,3,4,5)
dfPartial[[factorPos[[1]]]] %<>% factor()
dfPartial[[factorPos[[2]]]] %<>% factor()
dfPartial[[factorPos[[3]]]] %<>% factor()
dfPartial[[factorPos[[4]]]] %<>% factor()

# element 9 should be numeric
numPos = c(9)
dfPartial[[numPos]] %<>% as.numeric()
```

```{r}
dplyr::glimpse(dfPartial)
```

```{r log dfPartial}
# log
project$data$dfPartial$value <- dfPartial
```

  * (For foreign students) Translate element values from Chinese to English (method here only works on factor class)
```{r do the magic trick on levels}
levels(dfPartial$typeByFundingSource)
# change to English
levels(dfPartial$typeByFundingSource) <- c("public", "private")
levels(dfPartial$typeByFundingSource)

levels(dfPartial$day_night)
levels(dfPartial$day_night) <- c("part-time","day","night")
levels(dfPartial$day_night)
```

```{r log dfPartial again}
# log
project$data$dfPartial = list(
  value=dfPartial,
  description="dfPartial",
  source="enrollmentRate_department"
)
```


### Description

Introduce your data.

> Before we start, we rename features to English and keep only those we need for the next-step analysis. We also did the levels-swapping trick to translate Chinese element values into English quickly. A thorough class check was done as well. Wrong classes have been fixed.

  * data source and entity.
  
> The data is from Taiwan government open data platform. Each data entity is one tertiary education (高等教育）program. Each observation represents a certain program's freshmen registration rate.

```{r}
project$data$enrollmentRate_department$source |>
  browseURL()
```

> The data consists totaly 42,900 observations with 10 features

```{r}
project$data$dfPartial$value |>
  dplyr::glimpse()
```

> Among those features, each entity can be distinguished by the following four features: 

```{r}
names(project$data$dfPartial$value)[c(2:5)]
```
> For one example, the following programing is from a **public** university named **國立政治大學** under department **教育學系** for **day** time bachelor degress.

```{r}
project$data$dfPartial$value[c(2:5)][1,]
```

> Totally there are 8,290 different programs.

```{r}
project$data$dfPartial$value[c(2:5)] |> unique()
```

> Each program was tracked from  school year 106 (western calendar year 2017) to 110 (2021) -- if the program existed at the time.

```{r}
dfPartial = project$data$dfPartial$value
dfPartial$schoolYear |> unique()
```

```{r save-log}
# create data/ folder before you run the command
saveRDS(project, file="data/project.Rds") 
```

> Short summary:
> The data covers 8,290 tertiary educational programs from 2017 (school year 106) to 2021 (110) with total observations of 42,900 and ten features.

```{r save data 1}
saveRDS(project, file="data/project.Rds")
```


### Summary

```{r read data}
rm(list=ls()) # clean up global environment, emulate a new start of analysis on the next day
project = readRDS("data/project.Rds")
project$data$dfPartial$value -> dfPartial
```


For relevant features, 

  * any NA? How many as in percentage of the sample size? 

  * Distribution:
  
    * Discrete data: prop.table
    
    * Continuous data: range, mean, standard deviation
    
## Exploration

Any other things you discover in the data


## Conclusion


