---
title: "Final project"
date: "The Date"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, include = TRUE)
```

## Data 

  * log (for progress tracking purpose)
  
```{r}
project = list()
```
  


```{r}
# import data
library(readr)
enrollmentRate_department <- read_csv("data/學12-1.新生註冊率-以「系所」統計.csv")

# log 
project$data$enrollmentRate_department = list(
  value=enrollmentRate_department,
  description="enrollmentRate_department",
  source="https://data.gov.tw/dataset/26228"
)
```


### Checking

  * rename
  
```{r}
names(enrollmentRate_department)[c(1,2,5,7,8,10:14)]

names(enrollmentRate_department)[c(1,2,5,7,8,10:14)] <-  # notice 10:14
  c("schoolYear", "typeByFundingSource","school", "department","day_night", 
  "freshmenFresh","freshmenOld","freshmenFreshRegistered",
  "foreignStudentsRegistered", "netFreshmenRegistrationRate")

dfPartial = enrollmentRate_department[c(1,2,5,7,8,10:14)]
```

```{r}
dplyr::glimpse(dfPartial)
```

  * wrong class?

```{r}
library(magrittr)
dplyr::glimpse(dfPartial)
names(dfPartial)
```


```{r}
factorPos = c(2,3,4,5)
dfPartial[[factorPos[[1]]]] %<>% factor()
dfPartial[[factorPos[[2]]]] %<>% factor()
dfPartial[[factorPos[[3]]]] %<>% factor()
dfPartial[[factorPos[[4]]]] %<>% factor()

numPos = c(9)
dfPartial[[numPos]] %<>% as.numeric()
```

```{r}
dplyr::glimpse(dfPartial)
```

```{r}
# log
project$data$dfPartial$value <- dfPartial
```

  * (For foreign students) Translate element values from Chinese to English (method here only works on factor class)
```{r}
levels(dfPartial$typeByFundingSource)
# change to English
levels(dfPartial$typeByFundingSource) <- c("public", "private")
levels(dfPartial$typeByFundingSource)

levels(dfPartial$day_night)
levels(dfPartial$day_night) <- c("part-time","day","night")
levels(dfPartial$day_night)
```

```{r log-dfPartial}
# log
project$data$dfPartial = list(
  value=dfPartial,
  description="dfPartial",
  source="enrollmentRate_department"
)
```


### Description

Introduce your data.
```{r}
project$data$dfPartial$value |>
  dplyr::glimpse()
```


  * source: the data is from `r project$data$enrollmentRate_department$source`.
  
  * entity unit: a university program, like 國立政治大學教育學系日間部 is one entity. In total, there are 6,345 different programs.
```{r}
unique(dfPartial[c("school", "department", "day_night")])
```

  
  * sample size: Each program is tracked from school year 106 (western calendar year 2017) to 110 (2021) -- if the program existed at the time. In total, there 42,900 obs.

```{r}
dfPartial$schoolYear |> unique()
```

  
  * feature number: Each observation has the following features recorded.

```{r}
dfPartial |> names()
```

```{r save-log}
# create data/ folder before you run the command
saveRDS(project, file="data/project.Rds") 
```

### Summary

```{r read-log}
rm(list=ls())
project = readRDS("data/project.Rds")
project$data$dfPartial$value -> dfPartial
```


For relevant features, 

  * any NA? How many as in percentage of the sample size? 

  * Distribution:
  
    * Discrete data: prop.table
    
    * Continuous data: range, mean, standard deviation
    
## Exploration

Any other things you discover in the data


## Conclusion


